Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 approach.tex
started pass over sect 4

\subsection{Rendering}

We used a public rendering engine to create realistic rendering and
depth. We used CAD models that contain color, texture and material information such as
transparency and reflectance to make the rendering as much realistic
as possible to simulate the natural image statistics. The example rendering is
in Fig. \ref{fig:rendering}. We also extracted depth information as well while
rendering to use in the reconstruction.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which uses statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal,
\begin{equation}
\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X]) \label{eq:whitening}
\end{equation}
, we remove 2nd order correlation between all features. Since collecting covariance matrix for every template shape
 is expensive, \cite{Hariharan12} proposed easy way to synthesize
covariance $\Sigma$ from autocovariance $\Gamma$ and we followed their method to generate $\Sigma$

@@

\subsubsection{Whitening Synthesized Templates and Non-Zero WHO}
\label{sec:nzwho}
Our first innovation is `Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If the model is rendered over a natural image background, gradients
in the background will be incorporated into the discriminative template,
potentially causing false-positives if background elements exist in the query
image.

@@

\subsubsection{Fast Whitening using Conjugate Gradient}

The computational efficiency of our approach comes from using the iterative
Conjugate Gradient method to whiten a HOG template. Whitening the non-zero
feature vector $x'$ means a solving the system of linear equations, $\Sigma' w' = (x' -
\mu')$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elemination, which requires $O(n^3)$ time \scream{Citation for complexity}.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Realtime analysis of whitening. HOG means feature extraction time, WHO-Chol uses method of \cite{Hariharan12} and WHO-CG uses iterative Conjugate Gradient, and NZ-WHO-CG uses only non-zero cells and use Conjugate Gradient. (b) final speed of using NZ-WHO compare to WHO}
  \label{fig:whotime}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals from different method}
  \label{fig:whoresidual}
\end{figure}

@@

\subsection{Variations of WHO}

To validate our approach, we run an ensemble of NZ-WHO templates as a bank of exemplar detectors and compare its performance with other WHO variants. Since WHO variants uses HOG features, they are slower than HOG templates but WHO variants in general performs better than WHO by significant margin. However, WHO using Cholesky decomposition with Gaussian Elemination takes the most time yet performs worse than NZ-WHO which is two order of magnitude faster to generate.

>>>>

In this section, we describe our approach to generating 3D
exemplare-based templates on-the-fly. It is based on 3D CAD model
rendering (Sect.~\ref{sec:rendering}) followed by feature extraction
and applying a whitening procedure. Based on the original whitening
formulation (Sect.~\ref{sec:who}), we propose three novel extensions that
enable the application of whitening to the on-the-fly setting.
%
First, we adapt the whitening to the specific case of rendered
images (Sect.~\ref{sec:nzwho}). Second, we show how to speed up the
whitening by two orders of magnitudes for high-resolution templates
(Sect.~\ref{sec:fastwhiten}). And third, we improve the runtime of our
3D exemplar template detectors also at test time, by performing
convolutions in the frequency domain (Sect.~\ref{sec:fft}).

\subsection{Rendering}
\label{sec:rendering}
We used a public rendering engine to create realistic rendering and
depth. We used CAD models that contain color, texture and material information such as
transparency and reflectance to make the rendering as much realistic
as possible to simulate the natural image statistics. The example rendering is
in Fig. \ref{fig:rendering}. We also extracted depth information as well while
rendering to use in the reconstruction.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and training template detectors on-the-fly
is drawing from recent work, which we review briefly in this section.
%
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which uses statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal,
\begin{equation}
\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X]) \label{eq:whitening},
\end{equation}
we remove 2nd order correlation between all features. Since collecting covariance matrix for every template shape
 is expensive, \cite{Hariharan12} proposed easy way to synthesize
covariance $\Sigma$ from autocovariance $\Gamma$ and we followed their
method to generate $\Sigma$.

@@

\subsection{Whitening Synthesized Templates and Non-Zero WHO}
\label{sec:nzwho}
Our first innovation is `Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If the model is rendered over a natural image background, gradients
in the background will be incorporated into the discriminative template,
potentially causing false-positives if background elements exist in the query
image.

@@

\subsection{Fast Whitening using Conjugate Gradient}
\label{sec:fastwhiten}
Our second innovation lies in enhancing the computational efficiency
of the whitening procedure, making it applicable to an on-the-fly
rendering setting.
%
It originates from the insight that an iterative
Conjugate Gradient method can be used to whiten a HOG template, which
is much faster than the original procedure proposed
in~\cite{Hariharan12} based on decomposition.

Whitening the non-zero
feature vector $x'$ means a solving the system of linear equations, $\Sigma' w' = (x' -
\mu')$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elemination, which requires $O(n^3)$ time \scream{Citation for complexity}.

@@

\begin{figure}[t]
  \begin{center}
  \begin{tabular}{cc}
     \includegraphics[width=0.5\linewidth]{whotime} & 
     \includegraphics[width=0.5\linewidth]{speedup}\\
     (a) & (b) \\
 \end{tabular}
  \end{center}
  \caption{(a) Realtime analysis of whitening. HOG means feature extraction time, WHO-Chol uses method of \cite{Hariharan12} and WHO-CG uses iterative Conjugate Gradient, and NZ-WHO-CG uses only non-zero cells and use Conjugate Gradient. (b) final speed of using NZ-WHO compare to WHO}
  \label{fig:whotime}
\end{figure}
%
\begin{figure}[t]
  \centering
  \includegraphics[width=0.5\linewidth]{residual}
  \caption{Residuals from different method}
  \label{fig:whoresidual}
\end{figure}

@@

\subsection{Variations of WHO}
\label{sec:variations}
To validate our approach, we run an ensemble of NZ-WHO templates as a bank of exemplar detectors and compare its performance with other WHO variants. Since WHO variants uses HOG features, they are slower than HOG templates but WHO variants in general performs better than WHO by significant margin. However, WHO using Cholesky decomposition with Gaussian Elemination takes the most time yet performs worse than NZ-WHO which is two order of magnitude faster to generate.
